Hydras \& Co.

### 2022-04-11 doc/gaia-chapter.tex
small snippet change
corrections in gaia chapter

  We plan to make accessible the first kind of lemmas to \HydrasLib' users and the second kind to \gaia's. For this purpose, we write modules dedicated to the importation of definitions and lemmas from each of both libraries into the other one.
  Please note that by ``import'' we do not mean ``duplicate''
  nor ``repairing'' \HydrasLib' proofs.  Whenever possible, we proceed by rewriting steps over the statements and not the proofs. In a future version, it would be nice to have our proofs in the same context, after porting \HydrasLib' proofs under \gaia's
  definition, following  proof repair  techniques~\cite{ringer2021}.

@@

The modules are named {\texttt{theories/gaia/*.v}} and coqdocumented in \linebreak {\texttt{theories/html/gaia\_hydras.*.html}}
The main entry point is
\href{../theories/html/gaia_hydras.T1Bridge.html}{gaia\_hydras.T1Bridge}, which requires both Gaia's \texttt{ssete9} and a few modules of \texttt{theories/ordinals/Epsilon0}.
Other modules are dedicated to the adaptation of several notions of \HydrasLib to \mathcomp's and \gaia's vocabulary and structures.
By default, we give priority to \gaia's notation and vocabulary.

@@

Because of the common origin of \HydrasLib and \gaia, both libraries share the common name \texttt{T1} (data type of ordinal terms below $\epsilon_0$). As mentionned before, the bare name \texttt{T1} is given by priority to \gaia, and we prefix this name with an \texttt{h} (for ``hydras'') to refer to
\href{../theories/html/Epsilon0.T1.html}{hydras.Epsilon0.T1.T1}.

@@

There are a few exceptions to this convention.

@@

Please keep in mind that, in this module, we favour \gaia's notations over hydra's. A snippet like \texttt{(alpha < succ beta)} should be interpreted as \texttt{(T1lt alpha beta)}, where \texttt{T1lt} is defined in ~\href{https://github.com/coq-community/gaia/blob/master/theories/ssete9.v}{\texttt{gaia.ssete9.v}}.

@@

\subsubsection{Cancel lemmas}

@@

\inputsnippets{T1Bridge/eqRef}

\label{sect:gaia-compare-ref}
\inputsnippets{T1Bridge/compareRef}

\inputsnippets{T1Bridge/ltRef, T1Bridge/leRef}

\inputsnippets{T1Bridge/decideHLtRw}

@@

\inputsnippets{T1Bridge/SearchDemo}

\subsection{A type for well-formed ordinal terms}
\label{sect:E0-gaia-hydras}
For compatibility's sake, we add a clone of \texttt{hydra-battles}' type \texttt{E0} defined in \href{../theories/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}.

\section{E0 as an ordinal notation}
The module \href{../theories/html/gaia_hydras.T1Bridge.html}{gaia\_hydras.T1Bridge} defines an instance of class \texttt{ON E0lt compare}
(see Chapter~\ref{chap:ON}).

@@

We define this function by delegation to \HydrasLib'.
\vspace{4pt} 
\noindent.

From\href{../theories/html/gaia\_hydras.GHessenberg.html}{gaia\_hydras.GHessenberg}

@@

\inputsnippets{GHessenberg/oplusEquations}

@@

\inputsnippets{GHessenberg/oplusLemmasa}
\inputsnippets{GHessenberg/oplusLemmasb}
\inputsnippets{GHessenberg/oplusLemmasc}
\inputsnippets{GHessenberg/oplusLemmasd}
\inputsnippets{GHessenberg/oplusLemmase}

@@

\subsubsection{Transitions}

@@

\subsubsection{Paths}

@@

\inputsnippets{GPaths/Cor12}

@@

@@

The following definitions adapt \HydrasLib' to \ssreflect's 
definitions of $<$ and $\leq$ on \texttt{nat}.

\inputsnippets{T1Bridge/MonoDef}

>>>>

  We plan to make accessible the first kind of lemmas to \HydrasLib' users and the second kind to \gaia's. For this purpose, we write modules dedicated to the importation of definitions and lemmas from each of both libraries into the other one.
  Please note that by ``import'' we do not mean ``duplicate''
  nor ``repairing'' \HydrasLib' proofs.  Whenever possible, we proceed by rewriting steps over the statements and not the proofs. In a future version, it would be nice to have our proofs in the same context, after porting \HydrasLib' proofs under \gaia's
  definition, following  proof re-use or repair  techniques~\cite{Barthe2001, Magaud03a, ringer2021}.

@@

The modules are named {\texttt{theories/gaia/*.v}} and coqdocumented in \linebreak {\texttt{theories/html/gaia\_hydras.*.html}}
The main entry point is
\href{../theories/html/gaia_hydras.T1Bridge.html}{gaia\_hydras.T1Bridge}, which requires both Gaia's \texttt{ssete9} and a few modules of \texttt{theories/ordinals/Epsilon0}.
Other modules are dedicated to the adaptation of several notions developped in \HydrasLib to \mathcomp's and \gaia's vocabulary and structures.
By default, we give priority to \gaia's notation and vocabulary.

@@

Because of the common origin of \HydrasLib and \gaia, both libraries share the common name \texttt{T1} (data type of ordinal terms below $\epsilon_0$). As mentionned before, the bare name \texttt{T1} is given by priority to \gaia.
\HydrasLib' type \texttt{T1} may be called \texttt{hT1}
(\texttt{h} for ``hydras'') or simply \texttt{T1.T1}, \texttt{Epsilon0.T1.T1}, etc.

@@

The following notations are not defined in \gaia; we add them to \texttt{T1Bridge} for simplicity's sake.

@@

\begin{remark}
Please keep in mind that, in this module, we favour \gaia's notations over hydra's.
For instance, the mathematical inequality $42<\omega$ is 
expressed in \gaia's terms as follows:

\inputsnippets{HydraGaia_Examples/F42Omega}

The same inequality, with \HydrasLib vocabulary:

\inputsnippets{HydraGaia_Examples/F42Omegab}
\end{remark}

@@

@@

\inputsnippets{T1Bridge/eqRef}
Note that \HydrasLib' comparison on ordinal terms uses a
\stdpp-like trichotomy, hence the first two lemmas below:

\label{sect:gaia-compare-ref}
\inputsnippets{T1Bridge/compareRef}
\inputsnippets{T1Bridge/decideHLtRw}
\inputsnippets{T1Bridge/ltRef, T1Bridge/leRef}

@@

The following command lists us `gaia's lemmas (thanks to
\gaia's  \texttt{cantor\_scope}.
\inputsnippets{T1Bridge/SearchDemoa}

With \texttt{t1\_scope}:
\inputsnippets{T1Bridge/SearchDemob}

@@

We define this function in \href{../theories/html/gaia\_hydras.GHessenberg.html}{gaia\_hydras.GHessenberg} by delegation to \HydrasLib'.

\vspace{4pt} 
\noindent.

@@

\inputsnippets{GHessenberg/oplusEquations}
\inputsnippets{GHessenberg/oplusEquationsb}
\inputsnippets{GHessenberg/oplusEquationsc}

\inputsnippets{GHessenberg/oplusNf}

@@

\inputsnippets{GHessenberg/oplusCoplusA}
\inputsnippets{GHessenberg/oplusLemmasa}
\inputsnippets{GHessenberg/oplusLemmasb}
\inputsnippets{GHessenberg/oplusLemmasc}
\inputsnippets{GHessenberg/oplusLemmasd}
\inputsnippets{GHessenberg/oplusLemmase}

@@

\subsubsection{Transitions and paths}
Let us consider a kind of transition system, the states of which are ordinals below $\epsilon_0$. The elementary transitions (``small steps'') are jumps from an ordinal $\alpha$ to $\canonseq{\alpha}{i}$, for some $i\not=0$.

@@

Paths (sequences of transitions) are defined by delegation to
\href{../theories/html/hydras.Epsilon0.Paths.html}%
{\texttt{hydras.Epsilon0.Paths}}.

@@

\inputsnippets{GPaths/KSThm24}
\inputsnippets{GPaths/Cor12}

@@

\section{Rapidly growing arithmetic functions}
In this section, we consider a few families of arithmetic functions, indexed by ordinal numbers below $\epsilon_0$. In \HydrasLib, these functions are defined by transfinite induction, using the \texttt{coq-equations} plug-in~\cite{sozeau:hal-01671777}. For that purpose, we defined a sigma-type
\texttt{hydras.Epsilon0.E0.E0} of ordinal-terms in Cantor normal form.

These families share some monotony properties, also re-defined in
\href{../theories/html/gaia_hydras.T1Bridge.html}{gaia\_hydras.T1Bridge}:

\inputsnippets{T1Bridge/MonoDef}
\subsection{A type for well-formed ordinal terms}
\label{sect:E0-gaia-hydras}
For compatibility's sake, we add a clone of \texttt{hydra-battles}' type \texttt{E0} defined in \href{../theories/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}.

\subsection{E0 as an ordinal notation}
The module \href{../theories/html/gaia_hydras.T1Bridge.html}{gaia\_hydras.T1Bridge} defines an instance of class \texttt{ON E0lt compare}
(see Chapter~\ref{chap:ON}).

@@

### 2022-04-11 doc/part-hydras.tex
minor corrections
corrections in gaia chapter

\paragraph*{\gaiasign}
  In \texttt{gaia.ssete9}, the $n$-th finite ordinal is also written
  \texttt{$\backslash${F}\,$n$}. There is no coercion in \gaia from \texttt{nat} to finite ordinals.

@@

    In Module   \href{https://github.com/coq-community/hydra-battles/blob/master/theories/ordinals/OrdinalNotations/OmegaOmega.v}{theories/ordinals/OrdinalNotations/OmegaOmega.v},
    we represent ordinals below $\omega^\omega$ by lists of pairs of natural numbers (with the same coefficient shift as in \texttt{T1}).
    For instance, the ordinal $\omega^4\times 10 + \omega^3 + \omega+ 5$ is represented by the list \texttt{(4,9)::(3,0)::(1,0)::(0,4)::nil}.

>>>>

\paragraph*{\gaiasign}
  In \texttt{gaia.ssete9}, the $n$-th finite ordinal is also written
  \texttt{$\backslash${F}\,$n$}. There is no coercion in \gaia from \texttt{nat} to \texttt{T1}.

@@

    In Module   \href{https://github.com/coq-community/hydra-battles/blob/master/theories/
      ordinals/OrdinalNotations/OmegaOmega.v}{theories/ordinals/OrdinalNotations/OmegaOmega.v},
    we represent ordinals below $\omega^\omega$ by lists of pairs of natural numbers (with the same coefficient shift as in \texttt{T1}).
    For instance, the ordinal $\omega^4\times 10 + \omega^3 + \omega+ 5$ is represented by the list \texttt{(4,9)::(3,0)::(1,0)::(0,4)::nil}.
